---
title: "SVM Classifier"
author: "Siddharth.S.Chandran (18BCE1003)"
date: "26/03/2021"
output:
  pdf_document: default
  html_document: default
---

1. Read the dataset and find the structure of it 
```{r}
library(readr)
data<-read_csv("C:/Users/Siddharth.S.Chandran/Desktop/diabetes.csv")
str(data)
```

Here we can see the outcome variable as categorical variable. We encode this features as a factor

```{r}
data$Outcome = factor(data$Outcome, levels = c(0,1))
```

We divide the dataset into training and testing
```{r}
library(caTools)
set.seed(123)
split = sample.split(data$Outcome, SplitRatio = 0.75)
training_set = subset(data, split == TRUE)
test_set = subset(data, split == FALSE)
head(training_set)

```

The next step is to normalize the features of the training and testing data. This is done with the help of feature scaling. 
```{r}
cols<-c('Age', 'DiabetesPedigreeFunction','BMI', 'Insulin', 'SkinThickness','BloodPressure', 'Glucose', 'Pregnancies')
training_set[cols] = scale(training_set[cols])
test_set[cols] = scale(test_set[cols])
```

Now we improve the predictive accuracy of the algorithm. After scaling the features we proceed to fitting the SVM classifier data to the training set

Linear kernel

```{r}
library(e1071)
classifier = svm(formula = Outcome ~ .,
                 data = training_set,
                 type = 'C-classification',
                 kernel = 'linear')
summary(classifier)
```

The next line runs the classifier on the training set and test set so that the predictions can be made. 

```{r}
library(caret)
cols<-c('Age', 'DiabetesPedigreeFunction','BMI', 'Insulin', 'SkinThickness','BloodPressure', 'Glucose', 'Pregnancies')
y_pred = predict(classifier, newdata = test_set[cols])
y_train_pred = predict(classifier, newdata = training_set[cols])

```

To find the accuracy of the model we will find the confusion matrix
```{r}
cm1<-confusionMatrix(y_pred, test_set$Outcome)
cm1
cm2<-confusionMatrix(y_train_pred, training_set$Outcome)
cm2
```


Accruacy for the testing set -> 74.48%
Accuaracy for the training set -> 78.65% 

Radial kernel 
```{r}
library(e1071)
classifier = svm(formula = Outcome ~ .,
                 data = training_set,
                 type = 'C-classification',
                 kernel = 'radial')
summary(classifier)
```

The next line runs the classifier on the training set and test set so that the predictions can be made. 

```{r}
library(caret)
cols<-c('Age', 'DiabetesPedigreeFunction','BMI', 'Insulin', 'SkinThickness','BloodPressure', 'Glucose', 'Pregnancies')
y_pred = predict(classifier, newdata = test_set[cols])
y_train_pred = predict(classifier, newdata = training_set[cols])

```

To find the accuracy of the model we will find the confusion matrix
```{r}
cm1<-confusionMatrix(y_pred, test_set$Outcome)
cm1
cm2<-confusionMatrix(y_train_pred, training_set$Outcome)
cm2
```
Testing set accuracy -> 75% 
Training set accuracy -> 81.6%

Gaussian kernel 

